---
title: "Homework 3: Submission 1"
author: "Mathew Kang"
execute:
  echo: false
  warning: false
format: 
  pdf:
    mainfont: "Arial"
    geometry:
      - margin=1in
---

```{r}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata, scales, kableExtra, knitr, broom, here)

# Load dataset
final_ma <- read_csv(here::here("data", "output", "final_data.csv"))

```

# Question 1

```{r}
# Create summary table ----------------------------------------------------
q1_summary <- final_ma %>%
  group_by(year) %>%
  summarize(
    `Avg. Star Rating` = mean(as.numeric(Star_Rating), na.rm = TRUE),
    `Avg. Enrollment` = mean(as.numeric(avg_enrollment), na.rm = TRUE),
    `Avg. Market Share` = mean(as.numeric(avg_enrollment) / as.numeric(avg_enrolled), na.rm = TRUE),
    `Total Plans` = n_distinct(contractid, planid)
  ) %>%
  ungroup()


# Visualize table ---------------------------------------------------------
knitr::kable(q1_summary, 
             digits = 2, 
             caption = "Medicare Advantage Market Summary (2010-2015)",
             col.names = c("Year", "Mean Star Rating", "Mean Enrollment", "Mean Market Share", "Plan Count")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

# Question 2

```{r}
# Filter star ratings and create summary table ----------------------------------------------------
q2_nostar <- final_ma %>%
  filter(is.na(Star_Rating)) %>%
  group_by(year) %>%
  summarize(
    `Avg. Enrollment` = mean(as.numeric(avg_enrollment), na.rm = TRUE),
    `Avg. Market Share` = mean(as.numeric(avg_enrollment) / as.numeric(avg_enrolled), na.rm = TRUE),
    `Total Plans` = n_distinct(contractid, planid)
  ) %>%
  ungroup()


# Visualize table ---------------------------------------------------------
knitr::kable(q2_nostar, 
             digits = 2, 
             caption = "Medicare Advantage Market Summary Without Star Ratings (2010-2015)",
             col.names = c("Year", "Mean Enrollment", "Mean Market Share", "Plan Count")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

# Question 3

```{r}
# Filter for years 2010, 2012, and 2015 -----------------------------------
star_dist <- final_ma %>%
  filter(year %in% c(2010, 2012, 2015)) %>%
  filter(!is.na(Star_Rating)) %>%
  mutate(Star_Rating = as.factor(Star_Rating))


# Create bar graph --------------------------------------------------------
ggplot(star_dist, aes(x = Star_Rating)) +
  geom_bar(fill = "darkblue", color = "white") +
  facet_wrap(~year, ncol = 3) +
  labs(title = "Distribution of Medicare Advantage Star Ratings",
       x = "Star Rating",
       y = "Count of Plans") +
  theme_minimal()
```

# Question 4

```{r}
final_ma <- final_ma %>%
  mutate(
    market_share = as.numeric(avg_enrollment) / as.numeric(avg_eligibles),
    star_3 = if_else(as.numeric(Star_Rating) == 3, 1, 0),
    star_3.5 = if_else(as.numeric(Star_Rating) == 3.5, 1, 0),
    star_4 = if_else(as.numeric(Star_Rating) == 4, 1, 0),
    star_4.5_plus = if_else(as.numeric(Star_Rating) >= 4.5, 1, 0)
  )

q4_regression <- map_dfr(2010:2015, function(y) {
  model <- lm(market_share ~ star_3 + star_3.5 + star_4 + star_4.5_plus,
              data = filter(final_ma, year == y))
  tidy(model) %>%
    mutate(year = y)
}) %>%
  select(Rating = term, estimate, year) %>%
  pivot_wider(names_from = year, values_from = estimate)

q4_regression <- q4_regression %>%
  mutate(Rating = case_when(
    Rating == "(Intercept)" ~ "Excluded (≤ 2.5 Stars)",
    Rating == "star_3" ~ "3 Stars",
    Rating == "star_3.5" ~ "3.5 Stars",
    Rating == "star_4" ~ "4 Stars",
    Rating == "star_4.5_plus" ~ "≥ 4.5 Stars"
  ))

knitr::kable(q4_regression, 
             digits = 4) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

# Question 5

```{r}

quality <- c(
  "breastcancer_screen",
  "rectalcancer_screen",
  "cv_diab_cholscreen",
  "glaucoma_test",
  "monitoring",
  "flu_vaccine",
  "pn_vaccine",
  "physical_health",
  "mental_health",
  "osteo_test",
  "physical_monitor",
  "primaryaccess",
  "osteo_manage",
  "diab_healthy",
  "bloodpressure",
  "ra_manage",
  "copd_test",
  "bladder",
  "falling",
  "nodelays",
  "doctor_communicate",
  "carequickly",
  "customer_service",
  "overallrating_care",
  "overallrating_plan",
  "complaints_plan",
  "appeals_timely",
  "appeals_review",
  "leave_plan",
  "audit_problems",
  "hold_times",
  "info_accuracy",
  "ttyt_available"
)

rdup_2010 <- final_ma %>%
  filter(year == 2010) %>%
  rowwise() %>%
  mutate(
    raw_score = mean(c_across(all_of(quality)), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    rounded_rating = case_when(
      raw_score >= 2.75 & partc_score == 3.0 ~ "3.0 Stars",
      raw_score >= 3.25 & partc_score == 3.5 ~ "3.5 Stars",
      raw_score >= 3.75 & partc_score == 4.0 ~ "4.0 Stars",
      raw_score >= 4.25 & partc_score == 4.5 ~ "4.5 Stars",
      raw_score >= 4.75 & partc_score == 5.0 ~ "5.0 Stars",
      TRUE ~ NA_character_
    )
  ) %>% 
  filter(!is.na(rounded_rating)) %>%
  count(rounded_rating) %>%
  rename(`Star Rating` = rounded_rating, `Count` = n)

knitr::kable(rdup_2010,
             caption = "Plans Rounded Up Into Higher Star Ratings (2010)") %>%
    kable_styling(latex_options = c("striped", "hold_position"))
```

# Question 6

```{r}
rd1_2010 <- final_ma %>%
  filter(year == 2010) %>%
  rowwise() %>%
  mutate(
    raw_score = mean(c_across(all_of(quality)), na.rm = TRUE),
    market_share = as.numeric(avg_enrollment) / as.numeric(avg_enrolled)
  ) %>%
  ungroup() %>%
  filter(!is.na(raw_score), Star_Rating %in% c(2.5, 3)) %>%
  mutate(
    score = raw_score - 2.75,         
    treated = (score >= 0),              
    bandwidth = (score >= -0.125 & score <= 0.125), 
    score_treated = score * treated
  )

rd2_2010 <- final_ma %>%
  filter(year == 2010) %>%
  rowwise() %>%
  mutate(
    raw_score = mean(c_across(all_of(quality)), na.rm = TRUE),
    market_share = as.numeric(avg_enrollment) / as.numeric(avg_enrolled)
  ) %>%
  ungroup() %>%
  filter(!is.na(raw_score), Star_Rating %in% c(3, 3.5)) %>%
  mutate(
    score = raw_score - 3.25,         
    treated = (score >= 0),              
    bandwidth = (score >= -0.125 & score <= 0.125), 
    score_treated = score * treated
  )

reg1 <- lm(market_share ~ score + treated + score_treated, 
               data = (rd1_2010 %>% filter(bandwidth == TRUE)))

reg2 <- lm(market_share ~ score + treated + score_treated, 
               data = (rd2_2010 %>% filter(bandwidth == TRUE)))

est1 <- as.numeric(reg1$coefficients["treatedTRUE"])
est2 <- as.numeric(reg2$coefficients["treatedTRUE"])

p1 <- summary(reg1)$coefficients["treatedTRUE", 4]
p2 <- summary(reg2)$coefficients["treatedTRUE", 4]

rd_comparison <- data.frame(
  Comparison = c("2.5 vs 3.0 Stars", "3.0 vs 3.5 Stars"),
  Cutoff = c(2.75, 3.25),
  ATE = c(est1, est2),
  PValue = c(p1, p2)
)

knitr::kable(rd_comparison, 
             digits = 5, 
             col.names = c("Comparison", "Cutoff", "RD Estimate (ATE)", "P-Value"),
             caption = "Table 5: RDD Estimates of Star Rating Impact (Bandwidth = 0.125)")
```

# Question 7

```{r}
bandwidths <- c(0.1, 0.12, 0.13, 0.14, 0.15)
sensitivity_results <- data.frame()

for (h in bandwidths) {
  
  reg1 <- lm(market_share ~ score + treated + score_treated, 
             data = (rd1_2010 %>% filter(abs(score) <= h)))

  reg2 <- lm(market_share ~ score + treated + score_treated, 
             data = (rd2_2010 %>% filter(abs(score) <= h)))

  sensitivity_results <- rbind(sensitivity_results, 
    data.frame(Bandwidth = h, Estimate = coef(reg1)["treatedTRUE"], Comparison = "2.5 vs 3.0 Stars"),
    data.frame(Bandwidth = h, Estimate = coef(reg2)["treatedTRUE"], Comparison = "3.0 vs 3.5 Stars")
  )
}

ggplot(sensitivity_results, aes(x = Bandwidth, y = Estimate, color = Comparison, group = Comparison)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(title = "Sensitivity of RD Estimates to Bandwidth",
       y = "RD Estimate (ATE)",
       x = "Bandwidth (h)") +
  scale_color_manual(values = c("darkblue", "darkorange"))
```
# Question 8

```{r}
library(ggplot2)

density_plot <- function(df, cutoff, title_text) {
  df <- df %>% 
    mutate(side = ifelse(raw_score < cutoff, "Left", "Right"))
  
  ggplot(df, aes(x = raw_score, fill = side, color = side)) +
    geom_density(alpha = 0.3, size = 1) +
    geom_vline(xintercept = cutoff, linetype = "dashed", color = "red", size = 1.2) +
    geom_rug(alpha = 0.1) +
    theme_minimal() +
    labs(
      title = title_text,
      x = "Raw Quality Score",
      y = "Market Share",
      fill = "Side of Cutoff",
      color = "Side of Cutoff"
    ) +
    coord_cartesian(xlim = c(cutoff - 0.25, cutoff + 0.25)) +
    theme(legend.position = "none")
}

plot_275 <- density_plot(rd1_2010, 2.75, "2.5 vs 3.0 Stars")
plot_325 <- density_plot(rd2_2010, 3.25, "3.0 vs 3.5 Stars")

print(plot_275)
print(plot_325)
```
# Question 9
```{r}
hmo_types <- c("HMO", "HMOPOS")

rd1_2010 <- rd1_2010 %>%
  mutate(
    is_hmo = if_else(plan_type %in% hmo_types, 1, 0),
    is_partd = if_else(partd == "Yes", 1, 0)
  )

rd2_2010 <- rd2_2010 %>%
    mutate(
    is_hmo = if_else(plan_type %in% hmo_types, 1, 0),
    is_partd = if_else(partd == "Yes", 1, 0)
  )

hmo_1 <- lm(is_hmo ~ score + treated + score_treated, 
                data = (rd1_2010 %>% filter(bandwidth == TRUE)))

hmo_2 <- lm(is_hmo ~ score + treated + score_treated, 
                data = (rd2_2010 %>% filter(bandwidth == TRUE)))

hmo_jump_1 <- coef(hmo_1)["treatedTRUE"]
hmo_p_1 <- summary(hmo_1)$coefficients["treatedTRUE", 4]

hmo_jump_2 <- coef(hmo_2)["treatedTRUE"]
hmo_p_2 <- summary(hmo_2)$coefficients["treatedTRUE", 4]

balance_table <- data.frame(
  Comparison = c("2.5 vs 3.0 Stars", "3.0 vs 3.5 Stars"),
  `HMO Difference` = c(hmo_jump_1, hmo_jump_2),
  `P-Value` = c(hmo_p_1, hmo_p_2)
)

knitr::kable(balance_table, digits = 5, caption = "Table 6: Plan Characteristics")
```
# Question 10
While question 9 returned an error, the other findings suggest that increased star ratings are associated with more enrollments, with plans exceeding the 2.75 star threshold experiencing, on average, about a 5% statistically significant increase in enrollments. Interestingly, plans that exceed the 3.25 star cutoff actually experienced, on average, about a 4% decrease in enrollments. This could suggest that plans at the 3.5-star level encounter different competitive pressures that outweigh the benefits of a marginal increase in ratings. While more information from question 9 is needed, enrollment changes may be driven by the star ratings themselves and not necessarily quality measures, given the sheer number of them that makes star ratings a simpler guide to decision-making for enrolling. 